---
title: "Session 6: Homework 3"
author: "Your name goes here"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: zenburn
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
---


```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)
```


```{r load-libraries, echo=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(ggthemes)
library(GGally)
library(readxl)
library(here)
library(skimr)
library(janitor)
library(broom)
library(tidyquant)
library(infer)
library(openintro)
library(tidyquant)
```

# Omega Group plc- Pay Discrimination


At the last board meeting of Omega Group Plc., the headquarters of a large multinational company, the issue was raised that women were being discriminated in the company, in the sense that the salaries were not the same for male and female executives. A quick analysis of a sample of 50 employees (of which 24 men and 26 women) revealed that the average salary for men was about 8,700 higher than for women. This seemed like a considerable difference, so it was decided that a further analysis of the company salaries was warranted. 

You are asked to carry out the analysis. The objective is to find out whether there is indeed a significant difference between the salaries of men and women, and whether the difference is due to discrimination or whether it is based on another, possibly valid, determining factor. 

## Loading the data


```{r load_omega_data}
omega <- read_csv(here::here("data", "omega.csv"))
glimpse(omega) # examine the data frame
skim(omega)
summarize(omega)
```

## Relationship Salary - Gender ?


```{r, confint_single_valiables}
# Summary Statistics of salary by gender
mosaic::favstats (salary ~ gender, data=omega)

# Dataframe with two rows (male-female) and having as columns gender, mean, SD, sample size, 
# the t-critical value, the standard error, the margin of error, 
# and the low/high endpoints of a 95% condifence interval

male <- omega %>% 
filter(gender == "male") %>% 
summarise(average_salary_male = mean(salary, na.rm=TRUE),
          SD_salary_male = sd(salary, na.rm=TRUE),
          count_male =n(),
          SE_male= SD_salary_male/sqrt(count_male),
          T_critical=qt(0.975,count_male-1),
          lower95_male=average_salary_male - T_critical*SE_male,
          upper95_male=average_salary_male + T_critical*SE_male
          ) 
male

female <- omega %>% 
filter(gender == "female") %>% 
summarise(average_salary_female = mean(salary, na.rm=TRUE),
          SD_salary_female = sd(salary, na.rm=TRUE),
          count_female =n(),
          SE_female= SD_salary_female/sqrt(count_female),
          T_critical=qt(0.975,count_female-1),
          lower95_female=average_salary_female - T_critical*SE_female,
          upper95_female=average_salary_female + T_critical*SE_female
          ) 
female

```
Without running the hypothesis test we can cannot say that average salary of males is significantly greater than average salary of females.

> What can you conclude from your analysis? A couple of sentences would be enough


```{r, hypothesis_testing}
# hypothesis testing using t.test() 
t.test(salary ~ gender, data = omega)

# hypothesis testing using infer package

set.seed(1234)

difference_salary <- omega %>% 
  specify(salary~gender) %>% 
  hypothesise(null = "independence") %>% 
  generate(reps=1000, type ="permute") %>% 
  calculate(stat = "diff in means", order = c("male", "female"))
difference_salary %>% visualize()+
  
  shade_p_value(obs_stat = difference_salary, direction = "two-sided")+

labs(x="Salary difference between males and females", y="no. of employees", title="Hypothesis test for difference in salaries based on gender")
  
  difference_salary %>% get_p_value(obs_stat = difference_salary, direction = "two_sided")

glimpse(difference_salary)

```
Our null hyphotesis that average salary of men minus the average salary of women is zero. But zero does not lie in our 95% confidence interval and hence our hyphothesis fail, stating with 95% confidence that average salary of men is greater than average salary of women.  

## Relationship Experience - Gender?


```{r, experience_stats}
# Summary Statistics of salary by gender
favstats (experience ~ gender, data=omega)

male_1 <- omega %>% 
filter(gender == "male") %>% 
summarise(average_experience_male = mean(experience, na.rm=TRUE),
          SD_experience_male = sd(experience, na.rm=TRUE),
          count_male =n(),
          SE_male= SD_experience_male/sqrt(count_male),
          T_critical=qt(0.975,count_male-1),
          lower95_male=average_experience_male - T_critical*SE_male,
          upper95_male=average_experience_male + T_critical*SE_male
          ) 
male_1

female_1 <- omega %>% 
filter(gender == "female") %>% 
summarise(average_experience_female = mean(experience, na.rm=TRUE),
          SD_experience_female = sd(experience, na.rm=TRUE),
          count_female =n(),
          SE_female= SD_experience_female/sqrt(count_female),
          T_critical=qt(0.975,count_female-1),
          lower95_female=average_experience_female - T_critical*SE_female,
          upper95_female=average_experience_female + T_critical*SE_female
          ) 
female_1
```

Based on this evidence, without running a hypothesis test, we cannot conclude that there is a significant difference between the experience of the male and female executives

## Relationship Salary - Experience 


```{r, salary_exp_scatter}

ggplot(omega,aes(x=experience,y=salary)) + 
  geom_smooth() + 
  geom_point()+
  theme_classic()+ 
  labs (
    title = "Relationship in Experience and Salary",
    y     = "Salary",
    x     = "Experience in years"  
  )          
  
  
```
The curve shows that salary of employee increases with increase in experience.

## Relationship Gender - Experience

```{r, salary_exp_hyphothesis}
t.test(experience ~ gender, data = omega)

# hypothesis testing using infer package

set.seed(1234)
difference_experience <- omega %>% 
  specify(experience~gender) %>% 
  hypothesise(null = "independence") %>% 
  generate(reps=1000, type ="permute") %>% 
  calculate(stat = "diff in means", order = c("male", "female"))
difference_experience %>% visualize() +

shade_p_value(obs_stat = difference_experience, direction = "two-sided")+

labs(x="Salary difference based on experience", title="Hypothesis test for difference in salaries based on experience")
  
  difference_experience %>% get_p_value(obs_stat = difference_experience, direction = "two_sided")

  glimpse(difference_experience)




```
Our null hyphotesis that average experience of men minus the average experience of women is zero. But zero does not lie in our 95% confidence interval and hence our hypothesis fail, stating with 95% confidence that average experience of men is statistically greater than average experience of women. 


## Check correlations between the data


```{r, ggpairs}
omega %>% 
  select(gender, experience, salary) %>% #order variables they will appear in ggpairs()
  ggpairs(aes(colour=gender, alpha = 0.3))+
  theme_bw()


```

> Look at the salary vs experience scatterplot. What can you infer from this plot? Explain in a couple of sentences

*Based on the salary vs experience scatterplot, we can state that there are a lot of women with zero or very less work experience compared to men who generally have higher work experience. Moreover men with more workexperience have higher salaries. Further we can also see that there is no direct relationship between salary of employee based on gender, as in cases where male and female employees have same work experience, any one of them could have had higher salary * 

